---
title: "Quick analysis"
author: "Emily Wapman"
format: html
editor: visual
---

# Introduction

## About the data

The data for this analysis were downloaded from the Arctic Data Center on 10/7/2. [Arctic Data Center](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X)

## Setup

Attach important packages

```{r}
library(readr)
library(here)
```


## Read in the data

```{r}
#| message: false
#| warning: false
bg_chem <- read_csv(here('data', 'BGchem2008data.csv'))
```

# Analysis

## Calculate summary statistics
```{r}
#| eval: false
#| echo: false

### print the column names
colnames(bg_chem)

### get the general structure
str(bg_chem)

### First six lines or 20 lines
head(bg_chem, 20)

### Get a summary of each column
summary(bg_chem)

### Unique values of a column
unique(bg_chem$Date)
```
Calculate the summary statistics for nitrate, nitrite, ammonia, and phosphate.
```{r}
nitrate <- mean(bg_chem$NO3)
nitrite <- mean(bg_chem$NO2)
amm <- mean(bg_chem$NH4)
phos <- mean(bg_chem$P)
```


## Calculate the mean Redfield ratio

Redfield ratio is the relatie proportion of nitrogen to phosphorous

```{r}
ratio <- (nitrate + nitrite + amm) / phos
```

The Redfield ratio for this data set is approximately `r round(ratio)`

## Plot
```{r}
#| echo: false
plot(bg_chem$P, bg_chem$NO2 + bg_chem$NO3 + bg_chem$NH4)
```

#Conclusion
