---
title: "R Practice: Tidy Data and Joins"
author: "Emily Wapman"
format: html
---

### Load libraries 
```{r}
library(readr)
library(dplyr)
library(here)
library(lubridate) # for bonus question
```

### Load data
```{r}
#| message: false
#| warning: false
bird_observations <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=53edaa7a0e083013d9bf20322db1780e")

surveys <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=b2466fa5cb5ed7ee1ea91398fc291c59")

sites <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=81bf72420e69077097fb0790dcdc63a6")

taxalist <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=58f863b7e3066e68536a9cacdc7bd58e")
```

### Get familiar with data
```{r}
head(bird_observations) #primary: species + site ID + speciesID foreign: speciesID
head(surveys) # primary: surveyID foreign: site ID
head(sites) # primary: siteID foreign: none
head(taxalist) # primary: species ID foreign: none

summary(bird_observations) 
summary(surveys)
summary(sites)
summary(taxalist)

glimpse(bird_observations)
glimpse(surveys)
glimpse(sites)
glimpse(taxalist)
```
### Create a subset of bird_observations with species id BHCO and RWBL from sites LI-W and NU-C
```{r}
birds_subset <- bird_observations %>% 
  filter(species_id %in% c("BHCO", "RWBL")) %>% 
  filter(site_id %in% c("LI-W", "NU-C"))

```

### Use Left Join to merge birds_subset with the tables sites 
```{r}
birds_left <- left_join(birds_subset, sites, by = "site_id")
```

### Use full join to merge birds_subset and sites
```{r}
birds_full <- full_join(birds_subset, sites, by = "site_id")
head(birds_full, 20)

```
### Use inner_join() to merge birds_subset and taxalist
```{r}
birds_inner <- inner_join(birds_subset, taxalist, by = "species_id")
```

### What if you had used left_join() for this?
```{r}
birds_left_check <- left_join(birds_subset, taxalist, by = "species_id")

```

### Practice Wrangling and Joining Data
```{r}
bird_observations <- bird_observations %>%
  rename(bird_obs_notes = notes)

```

### Create a subset with all of birds_observations, join taxalist, sites, surveys and limit to only survey_dat, common_name, park_name, bird_count, observer
```{r}
clean_bird_obs <- _join(bird_observations, taxalist, by = "species_id") %>% 
  left_join(bird_)
```


